\myTitle

### 2024-05-07 body-text-of-paper.tex
added "lack-of-recombination-map" discussion piece

\gscramble{} is conducive for analysis of numerous types of marker sets available from contemporary data generation 
techniques. As mentioned above, 
one key feature of \gscramble{} is the integration of recombination rates, thereby accounting for physical 
linkage which is expected among larger contemporary data sets. Furthermore, \gscramble{} performs marker 
segregation with computational efficiency---for the four different GSPs used with the invasive wild pig data set, 
the mean run time of each simulation with 14,233 SNP loci was 2.3 seconds (range: 1.9-5.7 $s$). 
While \gscramble{} is equally applicable to data sets with 
microhaplotypes \citep{baetscher2018microhaplotypes} or microsatellites \citep{zhan2017megasat}, 
it is not suited for genotype likelihood data (i.e. from low-coverage whole genome sequencing). 
Specifically, simulations conducted within \gscramble{} involve the segregation of gametes within the 
specified pedigrees and, therefore, requires genotypes to be resolved.

We envision \gscramble{} to be relevant for a diverse range of potential applications. In our 
provided empirical examples with trout and wild pig data, we demonstrate the basics of 
creating hybrid classes. In much the same manner, \gscramble{} has been used to evaluate the statistical power 
of hybrid classification used to differentiate wild boar hybrids from domestic pigs 
\citep{smyser2024probgen}. Specifically \citet{smyser2024probgen} used the segregation options in \gscramble{} 
 to preserve founder individuals' genotypes within populations in contrast to \textit{a priori} shuffling alleles 
within populations. This retains the subpopulation structure present within certain populations---thereby, better reflecting the 
actual admixture variation expected from wild boar and domestic pig hybrids. We think \gscramble{} 
may also be useful for assessing relatedness estimates using pedigrees of varying complexity. Here, it 
is important to note that each time a gamete is segregated in a GSP, its complement is also 
segregated---an example of antithetical variate sampling in Monte Carlo methods. Thus, in the 
example of relatedness, fixed specified founders should be repeatedly "mated" with one offspring 
sampled per iteration to avoid sampling of anticorrelated genotypes.

@@

Also, we need to discuss the fact that, because each time a gamete is segregated in the
pedigree, its complement is also segregated, this is an example of antithetical variate sampling
in Monte Carlo.

>>>>

\gscramble{} is conducive for analysis of numerous types of marker sets available from contemporary data generation 
techniques. As mentioned above, 
one key feature of \gscramble{} is the integration of recombination rates, thereby accounting for physical 
linkage which is expected among high density markers sets. Furthermore, \gscramble{} performs marker 
segregation with computational efficiency---for the four different GSPs used with the invasive wild pig data set, 
the mean run time of each simulation with 14,233 SNP loci was 2.3 seconds (range: 1.9-5.7 $s$). 
While \gscramble{} is equally applicable to data sets with 
microhaplotypes \citep{baetscher2018microhaplotypes} or microsatellites \citep{zhan2017megasat}, 
it is not suited for genotype likelihood data (i.e. from low-coverage whole genome sequencing). 
Specifically, simulations conducted within \gscramble{} involve the segregation of gametes within the 
specified pedigrees and, therefore, requires genotypes to be resolved.

We envision \gscramble{} to be relevant for a diverse range of potential applications. In our 
provided empirical examples with trout and wild pig data, we demonstrate the basics of 
creating hybrid classes. In much the same manner, \gscramble{} has been used to evaluate the statistical power 
of hybrid classification used to differentiate wild boar hybrids from domestic pigs 
\citep{smyser2024probgen}. Specifically \citet{smyser2024probgen} used the segregation options in \gscramble{} 
 to preserve founder individuals' genotypes in contrast to \textit{a priori} shuffling alleles 
within populations. This retains the subpopulation structure present within certain populations---thereby, better reflecting the 
actual admixture variation expected from wild boar and domestic pig hybrids. We think \gscramble{} 
may also be useful for assessing relatedness estimates using pedigrees of varying complexity. Here, it 
is important to note that each time a gamete is segregated in a GSP, its complement is also 
segregated---an example of antithetical variate sampling in Monte Carlo methods. Thus, in the 
example of relatedness, fixed specified founders should be repeatedly "mated" with one offspring 
sampled per iteration to avoid sampling of anticorrelated genotypes.

Here, we have demonstrated the consequences of failing to consider physical linkage among loci 
given the availability of a recombination map. However, in the absence of a recombination map, 
researchers with a basic understanding of the genome size and the number of chromosomes in 
the study organism could still benefit from using gscramble to simulate assumed linkage. To do 
so, users of gscramble would need to randomly distribution markers across, and within, the 
known number of chromosomes to produce a conceptual marker map (*.map within the PLINK 
file structure), and assume basic rates of recombination (e.g., 1 cM per 1 Mb). The resulting 
simulations from gscramble would produce admixture fractions that capture the expected 
variation from known linkage patterns, in contrast to the reduced variation produced from 
assuming independence (Figure 5). Characterizing uncertainty would provide caution for 
researchers to not over interpret conclusions drawn in the absence of a recombination map.

@@
